---
title: "amj_run_TERGM_tutorial_3.R"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 3: Updating Data


```{r pressure, echo=FALSE}
##===============================
## If the required libraries are not installed on your computer, run:
## >  install.packages(c('btergm','parallel','texreg'))
##-------------------------------
library(btergm)
library(parallel)
library(texreg)

data_dir <- 'C:\\Users\\T430\\Google Drive\\PhD\\Dissertation\\competition networks\\compnet-awareness-tutorial\\data'
```


Set parameters for the analysis and load the data into memory:
```{r params}
## analysis parameters
firm_i <- 'qualtrics'  ## focal firm
d <- 2                 ## ego network theshold (order)

## load RDS data file into memory as a list of networks
# data_file <- file.path(data_dir,sprintf('%s_d%s.rds',firm_i,d))
data_file <- file.path(data_dir, 'tutorial_d2_competition_network_sample.rds')
nets.all <- readRDS(data_file)
len <- length(nets.all)

## set number of time periods
nPeriods <- 8  ## any number from 3 to 11 is OK, but use 8 to compare results with example

## subset network periods
nets <- nets.all[(len-nPeriods+1):len]
```


```{r print_data1}
net <- nets[[1]]
print(net)
```

```{r print_data2}
dim(net[,])
print(net[1:4,1:4])
```


```{r combine_data_files}

## missing value strings to convert to <NA> type in imported data file
na.strings <- c('NA', 'na', '')

##====================================
## Add New Data from Owler Data Files
##------------------------------------

## new data directory name
ower_data_dir <- 'owler_data'

## new data directory
owler_dir <- file.path(data_dir,ower_data_dir)

## init new dataframes
vt <- data.frame()  #vertices
el <- data.frame()  #edge list

## loop over data files in data directory
for (file in dir(owler_dir, pattern = '\\.csv$')) {
  cat(sprintf('\n\ndata file %s\n', file))  ## echo progress
  
  ## load data
  full_file_path <- file.path(owler_dir, file)
  df <- read.csv(full_file_path, stringsAsFactors = F, na.strings = na.strings)
  
  ## append verices
  vt <- rbind(vt, df)
  
  ## select competitor column names of the form: "competitor_<number>"
  cols <- names(df)
  compcols <- cols[grep('competitor_\\d{1,}',x = cols)]  ## \\d{1,} is integer of 1+ digits 
  
  ## if no competitor data columns or missing company_name_unique column, 
  ## then skip to next data file
  if (length(compcols)==0 | !('company_name_unique' %in% names(df)))
    next
  
  ## loop over firms in data file
  for (i in 1:nrow(df)) {
    
    ## firm i 
    firm_i <- df[i, 'company_name_unique']

    ## select competitors of firm i 
    firm_i_comps <- unlist(df[i, compcols]) ## unlist from data.frame to vector
    
    ## skip if firm i has no company_name_unique
    if (is.na(firm_i))
      next
    ## skip rows with no competitors included
    if (all(sapply(firm_i_comps, is.na)))
      next
    
    ## loop over each  competitor j of firm i
    for (j in 1:length(firm_i_comps)) {
      
      comp_j <- unname(firm_i_comps[j])
      
      if (!is.na(comp_j)) {
        tmp_el <- data.frame(source=firm_i, target=comp_j, rank=j, weight=1)
        ## append competitor relation
        el <- rbind(el, tmp_el)          
      }
      
    }
    
    if (i %% 50 == 0) cat(sprintf('firm %s %s\n', i, firm_i))  ## echo progress
    
  }
  
}

```

The `target` column is not of the format like `company_name_unique` which must be fixed or else the firm names cannot be matched against the vertex dataframe to create the graph data frame.

```{r}
## check vertices
dim(vt)
head(vt)

## check edge list
dim(el)
head(el, 20)
```


We need to update `target` column by using the mapping `name`-->`company_name_unique` that we already have in the vertex dataframe.

```{r}

# company name to company_name_unique mapping
mapping <- vt[,c('name','company_name_unique')]
names(mapping) <- c('target','target_name_unique')

## merge the original edge list and company_name_unique mapping
el2 <- merge(el, mapping, by.x='target', by.y='target', all.x=T, all.y=F)

## replace original target column with new mapped target_name_unique
el2$target <- el2$target_name_unique
## finally remove the temporary target_name_unique column
el2$target_name_unique <- NULL

head(el2)

## write edge list to csv file
el_file <- file.path(data_dir, 'owler_edge_list.csv')
write.csv(el, file = el_file, row.names = F)

## write vertex list to csv file
vt_file <- file.path(data_dir, 'owler_vertex_list.csv')
write.csv(vt, file = vt_file, row.names = F)
  

```






















